sWGS validation report

### 2019-08-06 Paper.Rmd
...

#library(biomaRt)
#library(GenomicRanges)
#library(stringr)
#library(ggplot2)
library(ggrepel)
library(GGally)
library(pander)
library(reshape2)
library(gridExtra)
library(CoxHD)
#library(plyr)
#library(dplyr)

@@

```{r, echo=F, message=F, fig.width=5, fig.height=5}
pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

@@

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

@@

rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'discovery'

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
)

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
}))

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
}))

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

ptendo = ptendo[-which(is.na(ptendo$mean)),]

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
#roc.plot(rocPtA)
rocPtA$model = 'Per Patient Avg'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  #print(unique(pt$Hospital.Research.ID))
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

@@

      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }

  pt$nl = (pt$Block-1)*2

@@

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

@@

pcfit = lm(perc~mn, data=pred.confidence)
rsid = rstandard(pcfit)
pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'

@@

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
```

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR

@@

# early/late samples
#df = subset(globalCNdf, Patient == pt)
df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'OR','Prediction')], by.x='variable', by.y='Samplename')
df = arrange(df, chr, start, Endoscopy.Year)

## count arms differently than segs??
df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

@@

el = subset(df2, early | late)
el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'

p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

@@

```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )
  sharedcfs = intersect(names(lcf), ch$label)
  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

nzcoefs = lapply(nzcoefs, function(x) {
  as_tibble(x, rownames='coef')
})

cfs = unique(unlist(sapply(nzcoefs, function(x) x[['coef']])))
loo.coefs = data.frame(matrix(ncol=0,nrow=length(cfs)))
loo.coefs$coef = cfs
for (pt in names(nzcoefs)) {
  loo.coefs = dplyr::full_join(loo.coefs, nzcoefs[[pt]], by='coef')
}
colnames(loo.coefs)[-1] = c(1:length(nzcoefs))
loo.coefs[is.na(loo.coefs)] = 0

@@

# in this case jack.se is the SE of the variance 
# per coefficient
ch$jack.se = apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})
colnames(ch) = c('label','coef','jack.se')
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
  coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')

ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se*1.96, ymax=coef+jack.se*1.96)) + 
  coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')

@@

p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
ga = grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
ncol=2)
ga

@@

p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
riskLegend = get.legend(p1)
p1 = p1 + theme(legend.position = 'bottom')
riskLegendLong = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')

>>>>

#library(biomaRt)
#library(GenomicRanges)
#library(stringr)
#library(ggplot2)
library(ggrepel)
library(GGally)
library(pander)
library(reshape2)
library(gridExtra)
library(CoxHD)
library(RColorBrewer)
#library(plyr)
#library(dplyr)

library(OACUtils)

@@

```{r, echo=F, message=F, fig.width=5, fig.height=5}
#pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

@@

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

@@

#rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}

file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = pg.samp %>% dplyr::select(Status, Prediction)
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'discovery'

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
)

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best', transpose=T)
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best', transpose=T)
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, Path.ID) %>% arrange(desc(Endoscopy.Year), PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best', transpose=T)
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(Endoscopy.Year, PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == min(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best', transpose=T)
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

# Excluding the final endoscopy
ptendo = do.call(bind_rows, lapply(unique(pg.samp$Hospital.Research.ID), function(pt) {
  x = pg.samp %>% filter(Hospital.Research.ID == pt)  
  x %>% filter(Endoscopy.Year != max(x$Endoscopy.Year))
})) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise_at(vars(Prediction), funs(max, mean), na.rm=T)

ptendo = ptendo %>% filter(!is.na(mean))

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best', transpose=T)
rocPtM$model = 'Per Patient Max'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best', transpose=T))
#roc.plot(rocPtA)
rocPtA$model = 'Per Patient Avg'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  #pt = left_join(pt, info %>% dplyr::select(Hospital.Research.ID, Patient,Samplename,Block), by=c('Hospital.Research.ID','Patient','Samplename'))

  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup

@@

      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }
  pt$nl = (pt$Block-1)*2

@@

back = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)  ), transform.by.time, info=patient.info))
#back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

@@

pcfit = lm(perc~mn, data=pred.confidence)
rsid = rstandard(pcfit)
#pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05, 'Moderate', Risk)))

@@

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

back = left_join(back, pred.confidence %>% dplyr::select(quants, Risk), by='quants')

@@

unadjRR = ggplot(back, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR

@@

# early/late samples
#df = subset(globalCNdf, Patient == pt)
df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'RR','Prediction')], by.x='variable', by.y='Samplename')
df = arrange(df, chr, start, Endoscopy.Year)

## count arms differently than segs??
df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(RR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  ) %>% mutate(Patient = factor(Patient))
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

@@

el = df2 %>% filter(early | late)
el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'

p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

@@

```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}

list.files(cache.dir)

load(paste0(cache.dir, '/all.pt.alpha.Rdata'), verbose=T)
rm(plots, performance.at.1se, dysplasia.df, models, cvs, labels)

coefs.per.pt = data.frame(matrix(ncol=length(unique(pg.samp$Patient)), nrow=nrow(ch), dimnames=list( ch$label, unique(pg.samp$Hospital.Research.ID))))

for (pt in unique(pg.samp$Hospital.Research.ID)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )
  sharedcfs = intersect(names(lcf), ch$label)
  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

nzcoefs = purrr::map(nzcoefs, as_tibble, rownames='coef')

cfs = unique(unlist(sapply(nzcoefs, function(x) x[['coef']])))
loo.coefs = data.frame(matrix(ncol=0,nrow=length(cfs)))
loo.coefs$coef = cfs
for (pt in names(nzcoefs)) {
  loo.coefs = dplyr::full_join(loo.coefs, nzcoefs[[pt]], by='coef')
}
colnames(loo.coefs)[-1] = names(nzcoefs)
loo.coefs[is.na(loo.coefs)] = 0

@@

# in this case jack.se is the SE of the variance 
# per coefficient
ch$jack.se = apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})
colnames(ch) = c('label','coef','jack.se')
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
  coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')
```

@@

p = pred.tiles(back %>% filter(Patient == 6), colors=myPal, probs=T, ncol=1) + labs(title='')
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
ga = grid.arrange(
  pred.tiles(back %>% filter(Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, colors=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, colors=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
ncol=2)

@@

p1 = pred.tiles(back %>% filter(Status == 'NP'), colors=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
riskLegend = get.legend(p1)
p1 = p1 + theme(legend.position = 'bottom')
riskLegendLong = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(back %>% filter(Status == 'P'), probs=T, colors=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')

### 2019-08-06 Validation Report.Rmd
...

sheets = readxl::excel_sheets(val.file)[1:10]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
}))

@@

sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = Cohort[1]) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)

vt = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = Cohort[1]) %>% group_by(Batch, `Sample Type`, Cohort) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
)

@@

val %>% group_by(as.integer(RA),`Sample Type`, Collection) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
) %>% rename('as.integer(RA)' = 'Processing') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')

@@

oldp = val %>% ungroup %>% filter(Batch %in% c(9))

oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

@@

  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))

@@

new.np = val %>% ungroup %>% filter(Batch %in% c(10))

@@

Again, all passed QC though on manual QC I might have failed 2-3.  There were 3 samples that just squeaked in under the cutoff threshold.  But they did.

@@

if (file.exists('batch_10.txt')) {
  new.np.preds = read_tsv('batch_10.txt')
} else {

  segFiles = grep(paste(unique(new.np$`Patient ID`),collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '2_segObj',  full.names = T, recursive = T), value=T)

  info = readxl::read_xlsx(val.file, 'Batch 10') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))

  new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
    load(f)  
    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(new.np.preds, 'batch_10.txt')
}

@@

new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low'))
) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')

>>>>

sheets = readxl::excel_sheets(val.file)[1:12]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
}))

@@

sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)

vt = val %>% group_by(Batch) %>% dplyr::mutate(Cohort =  paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, `Sample Type`, Cohort) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
)

@@

val %>% group_by(as.integer(RA),`Sample Type`, Collection) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
) %>% rename('as.integer(RA)' = 'Processing RA') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')

@@

oldp = val %>% ungroup %>% filter(Batch %in% c(8:12) & Cohort == 'Old progressor scrolls')

oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  `n Samples` = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

sheets = readxl::excel_sheets(val.file)[9:12]
info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
    dplyr::filter(Status == 'P') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
}))

@@

#  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))

@@

new.np = val %>% ungroup %>% filter(Batch %in% c(10:12) & Status == 'NP')

@@

Again, most passed QC though on manual QC I might have failed 2-3 others.  There were 3 samples that just squeaked in under the cutoff threshold.

@@

sheets = readxl::excel_sheets(val.file)[10:12]
info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
    dplyr::filter(Status == 'NP' & !is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
}))

if (file.exists('val_np.txt')) {
  new.np.preds = read_tsv('val_np.txt')
} else {
  segFiles = grep(paste(new.np$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)

  new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
    print(f)
    load(f)  
    if (nrow(BarrettsProgressionRisk::sampleResiduals(segmented) %>% dplyr::filter(Pass)) <= 0) return()

    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(new.np.preds, 'val_np.txt')
}

@@

val.new.np.preds = left_join(val %>% filter(sample %in% new.np.preds$Sample) %>% select(`Hospital Research ID`, sample, varMAD_median, n.segs, Status),
          new.np.preds %>% select(Sample, Probability, matches('Risk')), by=c('sample'='Sample'))

with(val.new.np.preds, cor.test(varMAD_median, Probability))

val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'QC var' = signif( mean(varMAD_median), 2)
) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')
